---
title: "Code for generating an instance of iSEE for the MouseGastrulation data from Dahlin et al, 2018"
author:
- name: Federico Marini
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  - Center for Thrombosis and Hemostasis (CTH), Mainz
  email: marinif@uni-mainz.de
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This command needs to be run once

```{r}
BiocManager::install("MouseGastrulationData")
library("MouseGastrulationData")
```

## EmbryoAtlasData

Preprocessing steps previously performed

... see vignette

```{r}
AtlasSampleMetadata

sce <- EmbryoAtlasData(samples = 21)
sce


singlets <- which(!(colData(sce)$doublet | colData(sce)$stripped))
plot(
    x = reducedDim(sce, "umap")[singlets, 1],
    y = reducedDim(sce, "umap")[singlets, 2],
    col = EmbryoCelltypeColours[colData(sce)$celltype[singlets]],
    pch = 19,
    xaxt = "n", yaxt = "n",
    xlab = "UMAP1", ylab = "UMAP2"
)

library("scater")
plotReducedDim(sce, dimred = "umap", color_by = "celltype")
```

### Launching iSEE

In its default configuration...

```{r}
library("iSEE")
iSEE(sce, appTitle = "iSEE the Embryo Atlas data")
```

... and with some pre-configured panel setup

```{r}
# start, explore, export, restart!
initial <- list()

################################################################################
# Settings for Reduced dimension plot 1
################################################################################

initial[["ReducedDimensionPlot1"]] <- new("ReducedDimensionPlot", Type = "umap", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "stage", FacetColumnByColData = "stage", 
    ColorByColumnData = "celltype", ColorByFeatureNameAssay = "counts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "stage", 
    SizeByColumnData = "sample", TooltipColumnData = character(0), 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "Column data", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "ENSMUSG00000051951", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "cell_52466", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_52466", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ReducedDimensionPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 6L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Row data table 1
################################################################################

initial[["RowDataTable1"]] <- new("RowDataTable", Selected = "ENSMUSG00000025903", Search = "", 
    SearchColumns = c("", ""), HiddenColumns = character(0), 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(RowDataTable = 1L), PanelHeight = 500L, 
    PanelWidth = 3L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 1
################################################################################

initial[["FeatureAssayPlot1"]] <- new("FeatureAssayPlot", Assay = "counts", XAxis = "Column data", 
    XAxisColumnData = "celltype", XAxisFeatureName = "ENSMUSG00000051951", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "ENSMUSG00000025903", YAxisFeatureSource = "RowDataTable1", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "stage", 
    FacetColumnByColData = "stage", ColorByColumnData = "cell", 
    ColorByFeatureNameAssay = "counts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "stage", SizeByColumnData = "sample", 
    TooltipColumnData = character(0), FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "None", ColorByDefaultColor = "#000000", ColorByFeatureName = "ENSMUSG00000051951", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "cell_52466", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_52466", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(FeatureAssayPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

initial_embryo <- initial

initial_embryo

registerAppOptions(sce, color.maxlevels=30)

iSEE(sce,
     initial = initial_embryo,
     appTitle = "iSEE the Embryo Atlas data")
```



## Tal1 Chimera Data

Alsu's dataset for the miloDE manuscript

```{r}
library("SingleCellExperiment")
library("MouseGastrulationData")
sce_chimera <- MouseGastrulationData::Tal1ChimeraData()
# delete doublets and stripped
sce_chimera <- sce_chimera[, !sce_chimera$celltype.mapped %in% c("Doublet" , "Stripped")]
# delete row for tomato - technically it is not a feature and it is stored in coldata anyway
sce_chimera <- sce_chimera[!rownames(sce_chimera) == "tomato-td" , ]

sce_chimera
```

Adding normalized values and some TSNE/UMAP to make the summary also more visual

```{r}
sce_chimera <- logNormCounts(sce_chimera)

library(scater)
sce_chimera <- runTSNE(sce_chimera)
sce_chimera <- runUMAP(sce_chimera)
```

### Launching iSEE

```{r}
library("iSEE")
iSEE(sce_chimera, appTitle = "iSEE on the Tal1 dataset, used in the miloDE package manuscript")
```

... and maybe with a more complex config to show how this whole can start at once with a nice set of pre-configured panels:

```{r}
initial <- list()

################################################################################
# Settings for Reduced dimension plot 1
################################################################################

initial[["ReducedDimensionPlot1"]] <- new("ReducedDimensionPlot", Type = "TSNE", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "stage", FacetColumnByColData = "stage", 
    ColorByColumnData = "stage.mapped", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "stage", 
    SizeByColumnData = "sample", TooltipColumnData = character(0), 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "Column data", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "ENSMUSG00000051951", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "cell_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_1", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ReducedDimensionPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Row data table 1
################################################################################

initial[["RowDataTable1"]] <- new("RowDataTable", Selected = "ENSMUSG00000026414", Search = "Tnnt", 
    SearchColumns = c("", ""), HiddenColumns = character(0), 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(RowDataTable = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 1
################################################################################

initial[["FeatureAssayPlot1"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "stage.mapped", XAxisFeatureName = "ENSMUSG00000051951", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "ENSMUSG00000026414", YAxisFeatureSource = "RowDataTable1", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "tomato", 
    FacetColumnByColData = "stage", ColorByColumnData = "tomato", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "stage", SizeByColumnData = "sample", 
    TooltipColumnData = character(0), FacetRowBy = "Column data", 
    FacetColumnBy = "None", ColorBy = "Column data", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "ENSMUSG00000051951", ColorByFeatureSource = "---", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "cell_1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = FALSE, 
    VisualChoices = c("Color", "Facet"), ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_1", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(FeatureAssayPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 2
################################################################################

initial[["ReducedDimensionPlot2"]] <- new("ReducedDimensionPlot", Type = "TSNE", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "tomato", FacetColumnByColData = "tomato", 
    ColorByColumnData = "celltype.mapped", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "stage", 
    SizeByColumnData = "sample", TooltipColumnData = character(0), 
    FacetRowBy = "None", FacetColumnBy = "Column data", ColorBy = "Column data", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "ENSMUSG00000051951", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "cell_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(
        xmin = -25.052183846846, xmax = 10.598819991308, ymin = -31.764029484869, 
        ymax = -14.436715080257, panelvar1 = TRUE, coords_css = list(
            xmin = 330L, xmax = 450L, ymin = 307.21875, ymax = 394.21875), 
        coords_img = list(xmin = 660L, xmax = 900L, ymin = 614.4375, 
            ymax = 788.4375), img_css_ratio = list(x = 2L, y = 2L), 
        mapping = list(x = "X", y = "Y", colour = "ColorBy", 
            panelvar1 = "FacetColumn"), domain = list(left = -35.9548832378364, 
            right = 37.93736089117, bottom = -35.4392733237053, 
            top = 38.6383366883993), range = list(left = 586.603809931507, 
            right = 1084.04109589041, bottom = 825.344109589041, 
            top = 81.4605601178167), log = list(x = NULL, y = NULL), 
        direction = "xy", brushId = "ReducedDimensionPlot2_Brush", 
        outputId = "ReducedDimensionPlot2"), VisualBoxOpen = TRUE, 
    VisualChoices = c("Color", "Facet"), ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_1", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 2
################################################################################

initial[["FeatureAssayPlot2"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "tomato", XAxisFeatureName = "ENSMUSG00000051951", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "ENSMUSG00000034664", YAxisFeatureSource = "---", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "stage", 
    FacetColumnByColData = "stage", ColorByColumnData = "cell", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "stage", SizeByColumnData = "sample", 
    TooltipColumnData = character(0), FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "None", ColorByDefaultColor = "#000000", ColorByFeatureName = "ENSMUSG00000051951", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "cell_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "cell_1", FontSize = 1, LegendPointSize = 1, 
    LegendPosition = "Bottom", HoverInfo = TRUE, LabelCenters = FALSE, 
    LabelCentersBy = "stage", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 1
################################################################################

initial[["ComplexHeatmapPlot1"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "ENSMUSG00000034664\nENSMUSG00000028717\nENSMUSG00000026414\n", 
    ClusterRows = FALSE, ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "tomato", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = NA_real_, 
    UpperBound = NA_real_, AssayCenterRows = TRUE, AssayScaleRows = FALSE, 
    DivergentColormap = "blue < white < orange", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 11L, 4L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ComplexHeatmapPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "ReducedDimensionPlot2", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

initial_tal1

registerAppOptions(sce, color.maxlevels=40)

iSEE(sce_chimera,
     initial = initial_tal1,
     appTitle = "iSEE the Embryo Atlas data")
```


# Session info {-}

```{r}
sessionInfo()
```
