---
title: "Code for generating an instance of iSEE for the data of the Tabula Muris project"
author:
- name: Federico Marini
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  - Center for Thrombosis and Hemostasis (CTH), Mainz
  email: marinif@uni-mainz.de
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Compiled date**: `r Sys.Date()`

**Last edited**: 2018-12-03

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```



# Code for generating an instance of iSEE for the data of the Tabula Muris project

Paper: https://www.biorxiv.org/content/early/2018/03/29/237446

Data: The data reported in this article have been made available on GEO (accession code: [GSE109774](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE109774)), as well as from http://tabula-muris.ds.czbiohub.org, following the link on Figshare.

An even cleverer and more straightforward way to access and explore the data is to use the TabulaMurisData Bioconductor package, http://bioconductor.org/packages/TabulaMurisData/.

# Getting the TabulaMurisData package

```{r}
library(BiocManager)
BiocManager::install("TabulaMurisData")
library(TabulaMurisData)
```


```{r}
library(ExperimentHub)
eh <- ExperimentHub()
query(eh, "TabulaMurisData")
```

Now loading all packages one might need throughout this document

```{r}
library(AnnotationHub)
library(SingleCellExperiment)
library(scater)
library(scran)
library(iSEE)
```


# FACS-sorted SmartSeq data

```{r}
# tm_droplet <- TabulaMurisDroplet()
tm_smartseq <- TabulaMurisSmartSeq2()
```

Keeping in mind that some operations are quite memory-intensive...

```{r}
pryr::object_size(tm_smartseq)
rownames(tm_smartseq) # no need to add gene symbols, but we can try with ENSEMBL ids
```

## Adding gene level annotation

```{r}
ah <- AnnotationHub()
query(ah, c("92", "EnsDb","Mus Musculus"))
edb_v92_mouse <- ah[["AH60992"]]
rowData(tm_smartseq)
ens_ids <- mapIds(edb_v92_mouse, keys=rownames(tm_smartseq), keytype="SYMBOL", column="GENEID")
# rowData(tm_smartseq)$SYMBOL <- rownames(tm_smartseq)
rowData(tm_smartseq)$ENSEMBL <- ens_ids

# We also determine the chromosomal location for each gene.
location <- mapIds(edb_v92_mouse, keys=rowData(tm_smartseq)$Symbol,
                   column="SEQNAME", keytype="SYMBOL")
rowData(tm_smartseq)$CHR <- location
## e.g. ...
summary(location=="MT")
# ok, no mitochondrial genes
head(rowData(tm_smartseq))
```

## Quality control on the cells

```{r}
tm_smartseq <- scater::calculateQCMetrics(tm_smartseq)
hist(log10(tm_smartseq$total_counts), breaks=20, col="grey80")
hist(log10(tm_smartseq$total_features_by_counts), breaks=20, col="grey80")

ave.counts <- calcAverage(tm_smartseq, use_size_factors=FALSE)
rowData(tm_smartseq)$AveCount <- ave.counts
hist(log10(ave.counts), col="grey80")

# We also examine the top-most expressed genes.
# This contains ribosomal protein genes and other usual suspects.

# plotHighestExprs(tm_smartseq,10)

libsize.drop <- isOutlier(tm_smartseq$total_counts, nmads=3, type="lower", 
    log=TRUE) # batch = ?
feature.drop <- isOutlier(tm_smartseq$total_features_by_counts, nmads=3, type="lower", 
    log=TRUE) 
spike.drop <- isOutlier(tm_smartseq$pct_counts_ERCC, nmads=3, type="higher")

keep <- !(libsize.drop | feature.drop | spike.drop)
data.frame(ByLibSize=sum(libsize.drop), ByFeature=sum(feature.drop),
    BySpike=sum(spike.drop), Remaining=sum(keep))

dim(tm_smartseq)
tm_smartseq$PassQC <- keep
# saveRDS(tm_smartseq, file="...")
tm_smartseq <- tm_smartseq[,keep]
dim(tm_smartseq)

num.cells <- nexprs(tm_smartseq, byrow=TRUE)
smoothScatter(log10(ave.counts), num.cells, ylab="Number of cells", 
    xlab=expression(Log[10]~"average count"))

# keeping cells where at least one gene is expressed
to.keep <- num.cells > 0
tm_smartseq <- tm_smartseq[to.keep,]
summary(to.keep)
```

## Normalizing for cell-specific biases

```{r}
set.seed(42)

# clusters <- quickCluster(tm_smartseq, min.mean=1, method="igraph",use.ranks = TRUE)
# tm_smartseq <- computeSumFactors(tm_smartseq, cluster=clusters, min.mean=0.1)
tm_smartseq <- computeSumFactors(tm_smartseq)

summary(sizeFactors(tm_smartseq))
plot(tm_smartseq$total_counts/1e6, sizeFactors(tm_smartseq), log="xy")

# for the spike-ins
tm_smartseq <- computeSpikeFactors(tm_smartseq, type="ERCC", general.use=FALSE)


# backup plan
tm_smartseq <- normalize(tm_smartseq)
tm_smartseq <- scater::runPCA(tm_smartseq)
tm_smartseq <- scater::runTSNE(tm_smartseq)

saveRDS(tm_smartseq,"tm_smartseq.rds")

```

## Setting up the iSEE config

```{r}
options(iSEE.maxlevels = 85)

tmu <- tm_smartseq

rdd <- redDimPlotDefaults(tmu, 5)
rdd$Downsample <- TRUE
rdd$SampleRes <- 200
rdd$PointSize <- 0.2
rdd$Type <- 2
rdd$ColorBy <- "Column data"
rdd$ColorByColData <- "tissue"

cdd <- colDataPlotDefaults(tmu, 5)
cdd$Downsample <- TRUE
cdd$SampleRes <- 200
cdd$YAxis <- "tissue"
cdd$XAxis <- "Column data"
cdd$XAxisColData <- "mouse_id"
cdd$PointSize <- 0.2

fad <- featAssayPlotDefaults(tmu, 5)
fad$Downsample <- TRUE
fad$SampleRes <- 200
fad$YAxisFeatName <- 2942
fad$XAxis <- "Column data"
fad$XAxisColData <- "tissue"
fad$PointSize <- 0.2
fad$ColorBy <- "Column data"
fad$ColorByColData <- "tissue"

rdp <- rowDataPlotDefaults(tmu, 5)
rdp$Downsample <- TRUE
rdp$SampleRes <- 200
rdp$YAxis <- "pct_dropout_by_counts"
rdp$XAxis <- "Row data"
rdp$XAxisRowData <- "log10_mean_counts"

sap <- sampAssayPlotDefaults(tmu, 5)
sap$Downsample <- TRUE
sap$SampleRes <- 200
sap$XAxis <- "Sample name"
sap$XAxisSampName <- "A1.B000127.3_38_F.1.1"
```

```{r}
initialPanels = DataFrame(
    Name = c(sprintf("Reduced dimension plot %i", 1),
             sprintf("Feature assay plot %i", 1),
             sprintf("Sample assay plot %i", 1),
             sprintf("Column data plot %i", 1),
             sprintf("Row data plot %i", 1),
             sprintf("Column statistics table %i", 1),
             sprintf("Row statistics table %i", 1)),
    Width = c(4, 4, 4, 3, 3, 3, 3))

iSEE(tmu, 
     redDimArgs = rdd,
     colDataArgs = cdd, 
     featAssayArgs = fad,
     rowDataArgs = rdp,
     sampAssayArgs = sap,
     initialPanels = initialPanels)
```


### An alternative setup, adding some custom panels

```{r}
CUSTOM_DIFFEXP <- function(se, ri, ci, assay="logcounts") {
    ri <- ri$active
    if (is.null(ri)) { # ignoring saved gene selections for now.
        ri <- rownames(se)
    }
    if (is.null(ci$active) || length(ci$saved)==0L) {
        return(data.frame(row.names=character(0), LogFC=integer(0))) # dummy value.
    }

    assayMatrix <- assay(se, assay)[ri, , drop=FALSE]
    active <- Matrix::rowMeans(assayMatrix[,ci$active,drop=FALSE])

    lfcs <- vector("list", length(ci$saved))
    for (i in seq_along(lfcs)) {
        saved <- Matrix::rowMeans(assayMatrix[,ci$saved[[i]],drop=FALSE])
        lfcs[[i]] <- active - saved
    }
    names(lfcs) <- sprintf("LogFC/%i", seq_along(lfcs))

    output <- do.call(data.frame, lfcs)
    rownames(output) <- ri
    output
}

CUSTOM_HEAT <- function(se, ri, ci, assay="logcounts") {
    everything <- CUSTOM_DIFFEXP(se, ri, ci, assay=assay)
    if (nrow(everything) == 0L) {
        return(ggplot()) # empty ggplot if no genes reported.
    }

    everything <- as.matrix(everything)
    top <- head(order(Matrix::rowMeans(abs(everything)), decreasing=TRUE), 50)
    topFC <- everything[top, , drop=FALSE]
    dimnames(topFC) <- list(gene=rownames(topFC), contrast=colnames(topFC))
    dfFC <- reshape2::melt(topFC)
    ggplot(dfFC, aes(contrast, gene)) + geom_raster(aes(fill = value))
}
```

```{r}
# rowData(tmu)$mean_log <- Matrix::rowMeans(logcounts(tmu))
# rowData(tmu)$var_log <- apply(logcounts(tmu), 1, var)

rdd2 <- redDimPlotDefaults(tmu, 5)
rdd2$Downsample <- TRUE
rdd2$SampleRes <- 200
rdd2$PointSize <- 0.2
rdd2$Type <- 2
rdd2$ColorBy <- "Column data"
rdd2$ColorByColData <- "tissue"

rdp2 <- rowDataPlotDefaults(tmu, 5)
rdp2$Downsample <- TRUE
rdp2$SampleRes <- 200
rdp2$YAxis <- "pct_dropout_by_counts"
rdp2$XAxis <- "Row data"
rdp2$XAxisRowData <- "log10_total_counts"

cdp2 <- customDataPlotDefaults(tmu, 1)
cdp2$Function <- "CUSTOM_HEAT"
cdp2$Arguments <- "assay logcounts"
cdp2$ColumnSource <- "Reduced dimension plot 1"
cdp2$RowSource <- "Row data plot 1"

cst2 <- customStatTableDefaults(tmu, 1)
cst2$Function <- "CUSTOM_DIFFEXP"
cst2$Arguments <- "assay logcounts"
cst2$ColumnSource <- "Reduced dimension plot 1"
cst2$RowSource <- "Row data plot 1"
```



```{r}
initialPanels2 = DataFrame(
    Name = c(sprintf("Reduced dimension plot %i", 1),
             sprintf("Row data plot %i", 1),
             sprintf("Custom data plot %i", 1),
             sprintf("Custom statistics table %i", 1)),
    Width = c(3, 3, 3, 3))

iSEE(tmu, 
     redDimArgs = rdd2,
     rowDataArgs = rdp2,
     customDataArgs=cdp2, 
     customStatArgs=cst2,
     initialPanels = initialPanels2,
     customDataFun=list(CUSTOM_HEAT=CUSTOM_HEAT),
     customStatFun=list(CUSTOM_DIFFEXP=CUSTOM_DIFFEXP),
     customSendAll=TRUE)
```










```{r eval=FALSE}
# We perform some pre-clustering to break up obvious clusters.
clusters <- quickCluster(tm_smartseq, method="igraph", subset.row=ave>=0.1,
                         irlba.args=list(maxit=1000)) # for convergence.
table(clusters)

# We then use the deconvolution method to compute size factors for each cell.
tm_smartseq <- computeSumFactors(tm_smartseq, min.mean=0.1, cluster=clusters)
summary(sizeFactors(tm_smartseq))
# We can plot these against the library sizes to see how much of a difference it makes.
plot(tm_smartseq$total_counts, sizeFactors(tm_smartseq), log="xy")
# Note that some size factors are very small and negative.
# This represents cells that have so few expressed features that it is not possible to obtain a sensible size factor.
neg.sf <- sizeFactors(tm_smartseq)<0
summary(neg.sf)
# Instead, we replace the size factor with the (scaled) library size.
library(Matrix)
lib.sizes <- colSums(counts(tm_smartseq))
scaled.lib.sizes <- lib.sizes/mean(lib.sizes)
sizeFactors(tm_smartseq)[neg.sf] <- scaled.lib.sizes[neg.sf]
# Finally, we compute normalized log-expresion values.
tm_smartseq <- normalize(tm_smartseq)
saveRDS(tm_smartseq,"sce_tabulamuris_norm.RDS")

```



The compact version - probably not as clean as the full one

```{r}
tm_smartseq <- computeSumFactors(tm_smartseq) 
tm_smartseq <- normalize(tm_smartseq)
tm_smartseq <- runPCA(tm_smartseq)
set.seed(42)
tm_smartseq <- runTSNE(tm_smartseq, use_dimred="PCA", perplexity=30)

## still giving this error on MBP
#> tm_smartseq <- scran::computeSumFactors(tm_smartseq)
# Error in nlsModel(formula, mf, start, wts) : 
  # singular gradient matrix at initial parameter estimates
# In addition: There were 12 warnings (use warnings() to see them)
```

## Clustering with graph-based methods 

```{r eval=FALSE}
# We use the shared nearest neighbour method for clustering.
snn.gr <- buildSNNGraph(tm_smartseq, use.dimred="PCA")
clusters <- igraph::cluster_walktrap(snn.gr)
tm_smartseq$Cluster <- factor(clusters$membership)
table(tm_smartseq$Cluster)

# Plotting them out to verify separateness.
plotTSNE(tm_smartseq, colour_by="Cluster")

# Also examining their modularity scores.
# We look at the ratio of the observed and expected edge weights, as the raw modularity varies by orders of magnitudes across clusters.
cluster.mod <- clusterModularity(snn.gr, tm_smartseq$Cluster, get.values=TRUE)
log.ratio <- log2(cluster.mod$observed/cluster.mod$expected + 1)
library(pheatmap)
pheatmap(log.ratio, cluster_rows=FALSE, cluster_cols=FALSE,
         color=colorRampPalette(c("white", "blue"))(100))

# Marker gene detection
# Detecting marker genes for each cluster.
marker.out <- findMarkers(tm_smartseq, clusters=tm_smartseq$Cluster)

saveRDS(tm_smartseq,"sce_tabulamuris_processed_smartseq.RDS")
```






# 10X droplet-based data

See the other document.

# Session info {-}

```{r}
sessionInfo()
```


