# Code for generating an instance of iSEE for the data of the Tabula Muris project

Paper: https://www.biorxiv.org/content/early/2018/03/29/237446

Data: The data reported in this article have been made available on GEO (accession code: [GSE109774](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE109774)), as well as from http://tabula-muris.ds.czbiohub.org, following the link on Figshare.

An even cleverer and more straightforward way to access and explore the data is to use the TabulaMurisData Bioconductor package, https://github.com/csoneson/TabulaMurisData.

## Getting the TabulaMurisData package

```{r}
library(BiocManager)
BiocManager::install("TabulaMurisData")
library(TabulaMurisData)
```


```{r}
library(ExperimentHub)
eh <- ExperimentHub()
query(eh, "TabulaMurisData")
```

```{r}
tm_droplet <- TabulaMurisDroplet()
tm_smartseq <- TabulaMurisSmartSeq2()
```

Now loading all packages one might need throughout this document

```{r}
library(AnnotationHub)
library(SingleCellExperiment)
library(scater)
library(scran)
library(iSEE)

```


### FACS-sorted SmartSeq data

Keeping in mind that some operations are quite memory-intensive...

```{r}
pryr::object_size(tm_smartseq)
rownames(tm_smartseq) # no need to add gene symbols, but we can try with ENSEMBL ids
```









### 10X droplet-based data

```{r}

```

## Session info

```{r}
sessionInfo()
```
