# Code for generating an instance of iSEE for the data of the Tabula Muris project

Paper: https://www.biorxiv.org/content/early/2018/03/29/237446

Data: The data reported in this article have been made available on GEO (accession code: [GSE109774](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE109774)), as well as from http://tabula-muris.ds.czbiohub.org, following the link on Figshare.

An even cleverer and more straightforward way to access and explore the data is to use the TabulaMurisData Bioconductor package, https://github.com/csoneson/TabulaMurisData.

## Getting the TabulaMurisData package

```{r}
library(BiocManager)
BiocManager::install("TabulaMurisData")
library(TabulaMurisData)
```


```{r}
library(ExperimentHub)
eh <- ExperimentHub()
query(eh, "TabulaMurisData")
```

```{r}
tm_droplet <- TabulaMurisDroplet()
tm_smartseq <- TabulaMurisSmartSeq2()
```

Now loading all packages one might need throughout this document

```{r}
library(AnnotationHub)
library(SingleCellExperiment)
library(scater)
library(scran)
library(iSEE)

```


### FACS-sorted SmartSeq data

Keeping in mind that some operations are quite memory-intensive...

```{r}
pryr::object_size(tm_smartseq)
rownames(tm_smartseq) # no need to add gene symbols, but we can try with ENSEMBL ids
```

Adding gene level annotation

```{r}
ah <- AnnotationHub()
query(ah, c("92", "EnsDb","Mus Musculus"))
edb_v92_mouse <- ah[["AH60992"]]
rowData(tm_smartseq)
ens_ids <- mapIds(edb_v92_mouse, keys=rownames(tm_smartseq), keytype="SYMBOL", column="GENEID")
# rowData(tm_smartseq)$SYMBOL <- rownames(tm_smartseq)
rowData(tm_smartseq)$ENSEMBL <- ens_ids

# We also determine the chromosomal location for each gene.
location <- mapIds(edb_v92_mouse, keys=rowData(tm_smartseq)$Symbol,
                   column="SEQNAME", keytype="SYMBOL")
rowData(tm_smartseq)$CHR <- location
## e.g. ...
summary(location=="MT")

head(rowData(tm_smartseq))
```

Quality control on the cells

```{r}
tm_smartseq <- scatercalculateQCMetrics(tm_smartseq)
hist(log10(tm_smartseq$total_counts), breaks=20, col="grey80")
hist(log10(tm_smartseq$total_features_by_counts), breaks=20, col="grey80")

ave <- calcAverage(tm_smartseq)
rowData(tm_smartseq)$AveCount <- ave
hist(log10(ave), col="grey80")

# We also examine the top-most expressed genes.
# This contains ribosomal protein genes and other usual suspects.

plotHighestExprs(tm_smartseq,10)
```

Normalizing for cell-specific biases

```{r}
# We perform some pre-clustering to break up obvious clusters.
clusters <- quickCluster(tm_smartseq, method="igraph", subset.row=ave>=0.1,
                         irlba.args=list(maxit=1000)) # for convergence.
table(clusters)

# We then use the deconvolution method to compute size factors for each cell.
tm_smartseq <- computeSumFactors(tm_smartseq, min.mean=0.1, cluster=clusters)
summary(sizeFactors(tm_smartseq))
# We can plot these against the library sizes to see how much of a difference it makes.
plot(tm_smartseq$total_counts, sizeFactors(tm_smartseq), log="xy")
# Note that some size factors are very small and negative.
# This represents cells that have so few expressed features that it is not possible to obtain a sensible size factor.
neg.sf <- sizeFactors(tm_smartseq)<0
summary(neg.sf)
# Instead, we replace the size factor with the (scaled) library size.
library(Matrix)
lib.sizes <- colSums(counts(tm_smartseq))
scaled.lib.sizes <- lib.sizes/mean(lib.sizes)
sizeFactors(tm_smartseq)[neg.sf] <- scaled.lib.sizes[neg.sf]
# Finally, we compute normalized log-expresion values.
tm_smartseq <- normalize(tm_smartseq)
saveRDS(tm_smartseq,"sce_tabulamuris_norm.RDS")

```



The compact version - probably not as clean as the full one

```{r}
tm_smartseq <- computeSumFactors(tm_smartseq) 
tm_smartseq <- normalize(tm_smartseq)
tm_smartseq <- runPCA(tm_smartseq)
set.seed(42)
tm_smartseq <- runTSNE(tm_smartseq, use_dimred="PCA", perplexity=30)

#> tm_smartseq <- scran::computeSumFactors(tm_smartseq)
# Error in nlsModel(formula, mf, start, wts) : 
  # singular gradient matrix at initial parameter estimates
# In addition: There were 12 warnings (use warnings() to see them)

saveRDS(tm_smartseq,"sce_tabulamuris_processed.RDS")

```











### 10X droplet-based data

```{r}

```

## Session info

```{r}
sessionInfo()
```
