---
title: "Code for generating an instance of iSEE for the data of the Tabula Muris project"
author:
- name: Federico Marini
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  - Center for Thrombosis and Hemostasis (CTH), Mainz
  email: marinif@uni-mainz.de
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Compiled date**: `r Sys.Date()`

**Last edited**: 2018-12-03

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```



# Code for generating an instance of iSEE for the data of the Tabula Muris project

Paper: https://www.biorxiv.org/content/early/2018/03/29/237446

Data: The data reported in this article have been made available on GEO (accession code: [GSE109774](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE109774)), as well as from http://tabula-muris.ds.czbiohub.org, following the link on Figshare.

An even cleverer and more straightforward way to access and explore the data is to use the TabulaMurisData Bioconductor package, http://bioconductor.org/packages/TabulaMurisData/.

# Getting the TabulaMurisData package

```{r}
library(BiocManager)
BiocManager::install("TabulaMurisData")
library(TabulaMurisData)
```


```{r}
library(ExperimentHub)
eh <- ExperimentHub()
query(eh, "TabulaMurisData")
```

```{r}
tm_droplet <- TabulaMurisDroplet()
# tm_smartseq <- TabulaMurisSmartSeq2()
```

Now loading all packages one might need throughout this document

```{r}
library(AnnotationHub)
library(SingleCellExperiment)
library(scater)
library(scran)
library(iSEE)
```


### FACS-sorted SmartSeq data

See the other document



### 10X droplet-based data

```{r}

```

## Session info

```{r}
sessionInfo()
```


