---
title: "Code for generating an instance of iSEE for the HCA Tonsil Data atlas - epithelial subset"
author:
- name: Federico Marini
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  - Center for Thrombosis and Hemostasis (CTH), Mainz
  email: marinif@uni-mainz.de
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# I have one minute to convince you


```{r}
devtools::install_github("massonix/HCATonsilData")
library("HCATonsilData")

tda_epithelial <- HCATonsilData::HCATonsilData(cellType = "epithelial")
```

A default instance can be run with

```{r}
library("iSEE")
iSEE(tda_epithelial)
```

Then I would work on that, configure some panels, store the code to generate them, and have a fully fledged instance, as displayed below.


# I have some more time to tell you the whole thing

## Retrieving the data

MORE on the data itself and how to get it

## Loading packages

```{r}
library("HCATonsilData")

tda_epithelial <- HCATonsilData::HCATonsilData(cellType = "epithelial")

library("iSEE")
library("iSEEu")

iSEEu::modeEmpty(tda_epithelial)

sce <- tda_epithelial
```


## Configuring the app

Go by hand first, then use this chunk of code to programmatically re-do it

```{r}
initial <- list()

################################################################################
# Settings for Reduced dimension plot 1
################################################################################

initial[["ReducedDimensionPlot1"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "donor_id", FacetColumnByColData = "donor_id", 
    ColorByColumnData = "annotation_20220215", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "donor_id", 
    SizeByColumnData = "age", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Column data", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "AL627309.1", ColorByFeatureSource = "---", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(xmin = 2.807778927129, 
        xmax = 6.2410123197737, ymin = -6.1386057683565, ymax = -2.8362244982929, 
        coords_css = list(xmin = 361L, xmax = 450L, ymin = 285L, 
            ymax = 366L), coords_img = list(xmin = 722L, xmax = 900L, 
            ymin = 570L, ymax = 732L), img_css_ratio = list(x = 2L, 
            y = 2L), mapping = list(x = "X", y = "Y", colour = "ColorBy"), 
        domain = list(left = -9.83895980185942, right = 9.19284266167208, 
            bottom = -7.93223003035989, top = 7.79266554230246), 
        range = list(left = 66.3150417380137, right = 1053.04109589041, 
            bottom = 819.98715432363, top = 48.5943700124905), 
        log = list(x = NULL, y = NULL), direction = "xy", brushId = "ReducedDimensionPlot1_Brush", 
        outputId = "ReducedDimensionPlot1"), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = c(ReducedDimensionPlot = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Row data table 1
################################################################################

initial[["RowDataTable1"]] <- new("RowDataTable", Selected = "KRT13", Search = "KRT[0-9]", 
    SearchColumns = c("", "", ""), HiddenColumns = character(0), 
    VersionInfo = list(iSEE = structure(list(c(2L, 8L, 0L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(RowDataTable = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 1
################################################################################

initial[["FeatureAssayPlot1"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "annotation_20220215", XAxisFeatureName = "AL627309.1", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "KRT13", YAxisFeatureSource = "RowDataTable1", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "donor_id", 
    FacetColumnByColData = "donor_id", ColorByColumnData = "barcode", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "donor_id", SizeByColumnData = "age", 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "None", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "AL627309.1", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = FALSE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = c(FeatureAssayPlot = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "ReducedDimensionPlot1", DataBoxOpen = TRUE, 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 2
################################################################################

initial[["ReducedDimensionPlot2"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "donor_id", FacetColumnByColData = "donor_id", 
    ColorByColumnData = "barcode", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "donor_id", 
    SizeByColumnData = "age", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Feature name", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "KRT13", ColorByFeatureSource = "RowDataTable1", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Column data plot 1
################################################################################

initial[["ColumnDataPlot1"]] <- new("ColumnDataPlot", XAxis = "None", YAxis = "annotation_20220215", 
    XAxisColumnData = "barcode", FacetRowByColData = "donor_id", 
    FacetColumnByColData = "donor_id", ColorByColumnData = "barcode", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "donor_id", SizeByColumnData = "age", 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "None", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "AL627309.1", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(xmin = 0.47990067299861, 
        xmax = 1.5127251238526, ymin = 1.4789673594537, ymax = 2.4678098848756, 
        coords_css = list(xmin = 135.015625, xmax = 496.015625, 
            ymin = 325.171875, ymax = 397.171875), coords_img = list(
            xmin = 270.03125, xmax = 992.03125, ymin = 650.34375, 
            ymax = 794.34375), img_css_ratio = list(x = 2L, y = 2L), 
        mapping = list(x = "X", y = "Y", height = "2 * YWidth", 
            width = "2 * XWidth", group = "interaction(X, Y)"), 
        domain = list(left = 0.4, right = 1.6, bottom = 0.4, 
            top = 6.6, discrete_limits = list(x = "", y = list(
                "Basal cells", "Crypt", "FDCSP epithelium", "Outer surface", 
                "Surface epithelium", "VEGFA+"))), range = list(
            left = 214.176369863014, right = 1053.04109589041, 
            bottom = 951.468161386986, top = 48.5943700124905), 
        log = list(x = NULL, y = NULL), direction = "xy", brushId = "ColumnDataPlot1_Brush", 
        outputId = "ColumnDataPlot1"), VisualBoxOpen = FALSE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = c(ColumnDataPlot = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 1
################################################################################

initial[["ComplexHeatmapPlot1"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "KRT5\nKRT14\nKRT4\nKRT13\nKRT78\nKRT80", 
    ClusterRows = TRUE, ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "annotation_20220215", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = 0L, 
    UpperBound = 0L, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 8L, 0L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ComplexHeatmapPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 2
################################################################################

initial[["FeatureAssayPlot2"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "annotation_20220215", XAxisFeatureName = "AL627309.1", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "TMPRSS2", YAxisFeatureSource = "---", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "donor_id", 
    FacetColumnByColData = "donor_id", ColorByColumnData = "barcode", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "donor_id", SizeByColumnData = "age", 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "None", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "AL627309.1", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = FALSE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 3
################################################################################

initial[["FeatureAssayPlot3"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "annotation_20220215", XAxisFeatureName = "AL627309.1", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "ACE2", YAxisFeatureSource = "---", YAxisFeatureDynamicSource = FALSE, 
    FacetRowByColData = "donor_id", FacetColumnByColData = "donor_id", 
    ColorByColumnData = "barcode", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "donor_id", 
    SizeByColumnData = "age", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "None", ColorByDefaultColor = "#000000", ColorByFeatureName = "AL627309.1", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = FALSE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = 3L, PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Row data table 2
################################################################################

initial[["RowDataTable2"]] <- new("RowDataTable", Selected = "LCE3E", Search = "LCE3", SearchColumns = c("", 
"", ""), HiddenColumns = character(0), VersionInfo = list(iSEE = structure(list(
    c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 3
################################################################################

initial[["ReducedDimensionPlot3"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "donor_id", FacetColumnByColData = "donor_id", 
    ColorByColumnData = "barcode", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "donor_id", 
    SizeByColumnData = "age", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Feature name", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "LCE3E", ColorByFeatureSource = "RowDataTable2", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "bw94nf57_vm85woki_CTCCTTTCACACGGAA-1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "donor_id", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 8L, 0L)), class = c("package_version", "numeric_version"
    ))), PanelId = 3L, PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 2
################################################################################

initial[["ComplexHeatmapPlot2"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "CNFN\nLCE3A\nLCE3D\nLCE3E\nSPRR2D\nSPRR2E", 
    ClusterRows = FALSE, ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "annotation_20220215", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = 0L, 
    UpperBound = 0L, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 8L, 0L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 3
################################################################################

initial[["ComplexHeatmapPlot3"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "IL1B\nS100A6\nSPIB\nMARCKSL1", ClusterRows = FALSE, 
    ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "annotation_20220215", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = 0L, 
    UpperBound = 0L, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 8L, 0L)), class = c("package_version", 
    "numeric_version"))), PanelId = 3L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "ColumnDataPlot1", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 4
################################################################################

initial[["ComplexHeatmapPlot4"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "FDCSP\nKRTDAP\n", ClusterRows = FALSE, 
    ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "annotation_20220215", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = 0L, 
    UpperBound = 0L, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 8L, 0L)), class = c("package_version", 
    "numeric_version"))), PanelId = 4L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())
```


## Figures figures figures

Do some by hand, then extract the code

```{r}

```

## Adding a tour to tell things to people - they like to hear or read about the data!

```{r}
tour_epithelial <- read.delim("tour_tda_epithelial.txt", sep = ";", header = TRUE)
```


## Running the fully fledged instance of iSEE

This is just it ðŸ’£ ðŸ’£ ðŸ’£ ðŸ’£ ðŸ’£ 

(and this is the line that starts the app, fully configured)

```{r}
iSEE(sce, initial = initial, appTitle = "iSEE â¤ï¸ Tonsil Data Atlas - epithelial cells")
```

To start the app with the tour running, you can just run this command:

```{r}
iSEE(sce, initial = initial, appTitle = "iSEE â¤ï¸ Tonsil Data Atlas - epithelial cells",
     tour = tour_epithelial)
```



# Session info

```{r}
sessionInfo()
```



