---
title: "Code for generating an instance of iSEE for the TICA"
author:
- name: Federico Marini
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  - Center for Thrombosis and Hemostasis (CTH), Mainz
  email: marinif@uni-mainz.de
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# I have one minute to convince you

Retrieving the datasets from the Zenodo archive

https://zenodo.org/record/4263972#.X7BQJ9u1t25

```
wget -c https://zenodo.org/record/4263972/files/TICAtlas_downsampled_1000.h5ad?download=1
```

```{r}
library("zellkonverter")
sce_tica <- readH5AD("TICAtlas_downsampled_1000.h5ad")
library("iSEE")
iSEE(sce_tica)
```



# I have some more time to tell you the whole thing

## Retrieving the data

From the zenodo archive: https://zenodo.org/record/4263972#.X7BQJ9u1t25

(This is from the preprint version, with a nice downsampled file in AnnData format :) )

    TICAtlas.rds: an rds file containing a Seurat object with the whole Atlas (317111 cells, RNA and integrated assays, PCA and UMAP reductions)
    TICAtlas.h5ad: an h5ad file with the whole Atlas (317111 cells, RNA assay, PCA and UMAP)
    TICAtlas_RNA.rds: an rds file containing a Seurat object of the whole Atlas but only the RNA assay (317111 cells, UMAP embedding)
    TICAtlas_downsampled_1000.rds: an rds file containing a downsampled version of the Seurat object of the whole Atlas (24834 cells, RNA and integrated assay, PCA  and UMAP reductions)
    TICAtlas_downsampled_1000.h5ad: an rds file containing a downsampled version of the Seurat object of the whole Atlas (24834 cells, RNA assay, PCA  and UMAP reductions)
    TICAtlas_metadata.csv: a comma-separated text file with the metadata for each of the cells

```
wget -c https://zenodo.org/record/4263972/files/TICAtlas_RNA.rds?download=1
wget -c https://zenodo.org/record/4263972/files/TICAtlas_downsampled_1000.rds?download=1

wget -c https://zenodo.org/record/4263972/files/TICAtlas.h5ad?download=1
wget -c https://zenodo.org/record/4263972/files/TICAtlas_metadata.csv?download=1
wget -c https://zenodo.org/record/4263972/files/TICAtlas_downsampled_1000.h5ad?download=1
```

Extra info: 
https://github.com/Single-Cell-Genomics-Group-CNAG-CRG/Tumor-Immune-Cell-Atlas

## Loading packages

```{r}
library("zellkonverter")
sce_tica <- readH5AD("TICAtlas_downsampled_1000.h5ad")
library("iSEE")
library("iSEEu")

iSEEu::modeEmpty(sce_tica)
```

To make the loading quickest (and independent of `zellkonverter`) do the following:

```{r}
sce_tica
colData(sce_tica)
colnames(sce_tica)

reducedDimNames(sce_tica) <- "UMAP"
assayNames(sce_tica) <- "logcounts"
sce_tica

HDF5Array::saveHDF5SummarizedExperiment(sce_tica, dir = "see_tica", replace = TRUE)
sce <- HDF5Array::loadHDF5SummarizedExperiment("see_tica/")
scater::plotUMAP(sce, colour_by="cell_type")

# why are we even doing this?
microbenchmark::microbenchmark({sce <- HDF5Array::loadHDF5SummarizedExperiment("see_tica/")}, times = 10)
```


## Configuring the app

Go by hand first, then use this chunk of code to programmatically re-do it

```{r}
iSEEu::modeEmpty(sce)

initial <- list()

################################################################################
# Settings for Column data plot 1
################################################################################

initial[["ColumnDataPlot1"]] <- new("ColumnDataPlot", XAxis = "None", YAxis = "source", XAxisColumnData = "nCount_RNA", 
    FacetRowByColData = "gender", FacetColumnByColData = "gender", 
    ColorByColumnData = "nCount_RNA", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "gender", 
    SizeByColumnData = "nCount_RNA", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "None", ColorByDefaultColor = "#000000", ColorByFeatureName = "TSPAN6", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "X121270372616437_1_1_1_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "X121270372616437_1_1_1_1", FontSize = 1, 
    LegendPointSize = 1, LegendPosition = "Bottom", HoverInfo = TRUE, 
    LabelCenters = FALSE, LabelCentersBy = "gender", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ColumnDataPlot = 1L), PanelHeight = 500L, 
    PanelWidth = 3L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Column data plot 2
################################################################################

initial[["ColumnDataPlot2"]] <- new("ColumnDataPlot", XAxis = "Column data", YAxis = "source", 
    XAxisColumnData = "cell_type", FacetRowByColData = "gender", 
    FacetColumnByColData = "gender", ColorByColumnData = "source", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "gender", SizeByColumnData = "nCount_RNA", 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "Column data", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "TSPAN6", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "X121270372616437_1_1_1_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = TRUE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "X121270372616437_1_1_1_1", FontSize = 1, 
    LegendPointSize = 1, LegendPosition = "Bottom", HoverInfo = TRUE, 
    LabelCenters = FALSE, LabelCentersBy = "gender", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 3L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 1
################################################################################

initial[["ReducedDimensionPlot1"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "gender", FacetColumnByColData = "gender", 
    ColorByColumnData = "cell_type", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "gender", 
    SizeByColumnData = "nCount_RNA", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Column data", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "TSPAN6", ColorByFeatureSource = "---", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "X121270372616437_1_1_1_1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "X121270372616437_1_1_1_1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "gender", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 5L, 2L)), class = c("package_version", "numeric_version"
    ))), PanelId = c(ReducedDimensionPlot = 1L), PanelHeight = 500L, 
    PanelWidth = 3L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 2
################################################################################

initial[["ReducedDimensionPlot2"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "gender", FacetColumnByColData = "gender", 
    ColorByColumnData = "subtype", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "gender", 
    SizeByColumnData = "nCount_RNA", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Column data", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "TSPAN6", ColorByFeatureSource = "---", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "X121270372616437_1_1_1_1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "X121270372616437_1_1_1_1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "gender", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 5L, 2L)), class = c("package_version", "numeric_version"
    ))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 3L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Row data table 1
################################################################################

initial[["RowDataTable1"]] <- new("RowDataTable", Selected = "CD14", Search = "^CD1", SearchColumns = "", 
    HiddenColumns = character(0), VersionInfo = list(iSEE = structure(list(
        c(2L, 5L, 2L)), class = c("package_version", "numeric_version"
    ))), PanelId = c(RowDataTable = 1L), PanelHeight = 500L, 
    PanelWidth = 4L, SelectionBoxOpen = FALSE, RowSelectionSource = "---", 
    ColumnSelectionSource = "---", DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, 
    ColumnSelectionDynamicSource = FALSE, RowSelectionRestrict = FALSE, 
    ColumnSelectionRestrict = FALSE, SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 1
################################################################################

initial[["FeatureAssayPlot1"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "cell_type", XAxisFeatureName = "TSPAN6", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "CD14", YAxisFeatureSource = "RowDataTable1", 
    YAxisFeatureDynamicSource = FALSE, FacetRowByColData = "gender", 
    FacetColumnByColData = "gender", ColorByColumnData = "nCount_RNA", 
    ColorByFeatureNameAssay = "logcounts", ColorBySampleNameColor = "#FF0000", 
    ShapeByColumnData = "gender", SizeByColumnData = "nCount_RNA", 
    FacetRowBy = "None", FacetColumnBy = "None", ColorBy = "None", 
    ColorByDefaultColor = "#000000", ColorByFeatureName = "TSPAN6", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "X121270372616437_1_1_1_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "X121270372616437_1_1_1_1", FontSize = 1, 
    LegendPointSize = 1, LegendPosition = "Bottom", HoverInfo = TRUE, 
    LabelCenters = FALSE, LabelCentersBy = "gender", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(FeatureAssayPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 1
################################################################################

initial[["ComplexHeatmapPlot1"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "MS4A1\nCD79A\nCD83\nCD79B\nTCL1A\nIGKC\nLTB\nTCF7\nFOXP3\nBATF\nCREM\nSTMN1\nTUBB\nTYMS\nMX1\nIFI44L\nIFI6\nPRF1\nGZMB\nC1QA\nC1QC\nIL1B\nTSPAN13\nTCF4\nIRF7\n", 
    ClusterRows = TRUE, ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "subtype", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = NA_real_, 
    UpperBound = NA_real_, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = c(ComplexHeatmapPlot = 1L), 
    PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Reduced dimension plot 3
################################################################################

initial[["ReducedDimensionPlot3"]] <- new("ReducedDimensionPlot", Type = "UMAP", XAxis = 1L, YAxis = 2L, 
    FacetRowByColData = "gender", FacetColumnByColData = "gender", 
    ColorByColumnData = "cluster_kmeans_k6", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "gender", 
    SizeByColumnData = "nCount_RNA", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "Column data", ColorByDefaultColor = "#000000", 
    ColorByFeatureName = "TSPAN6", ColorByFeatureSource = "---", 
    ColorByFeatureDynamicSource = FALSE, ColorBySampleName = "X121270372616437_1_1_1_1", 
    ColorBySampleSource = "---", ColorBySampleDynamicSource = FALSE, 
    ShapeBy = "None", SizeBy = "None", SelectionAlpha = 0.1, 
    ZoomData = numeric(0), BrushData = list(), VisualBoxOpen = TRUE, 
    VisualChoices = "Color", ContourAdd = FALSE, ContourColor = "#0000FF", 
    PointSize = 1, PointAlpha = 1, Downsample = FALSE, DownsampleResolution = 200, 
    CustomLabels = FALSE, CustomLabelsText = "X121270372616437_1_1_1_1", 
    FontSize = 1, LegendPointSize = 1, LegendPosition = "Bottom", 
    HoverInfo = TRUE, LabelCenters = FALSE, LabelCentersBy = "gender", 
    LabelCentersColor = "#000000", VersionInfo = list(iSEE = structure(list(
        c(2L, 5L, 2L)), class = c("package_version", "numeric_version"
    ))), PanelId = 3L, PanelHeight = 500L, PanelWidth = 4L, SelectionBoxOpen = FALSE, 
    RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Feature assay plot 2
################################################################################

initial[["FeatureAssayPlot2"]] <- new("FeatureAssayPlot", Assay = "logcounts", XAxis = "Column data", 
    XAxisColumnData = "cell_type", XAxisFeatureName = "TSPAN6", 
    XAxisFeatureSource = "---", XAxisFeatureDynamicSource = FALSE, 
    YAxisFeatureName = "C1QC", YAxisFeatureSource = "---", YAxisFeatureDynamicSource = FALSE, 
    FacetRowByColData = "gender", FacetColumnByColData = "gender", 
    ColorByColumnData = "nCount_RNA", ColorByFeatureNameAssay = "logcounts", 
    ColorBySampleNameColor = "#FF0000", ShapeByColumnData = "gender", 
    SizeByColumnData = "nCount_RNA", FacetRowBy = "None", FacetColumnBy = "None", 
    ColorBy = "None", ColorByDefaultColor = "#000000", ColorByFeatureName = "TSPAN6", 
    ColorByFeatureSource = "---", ColorByFeatureDynamicSource = FALSE, 
    ColorBySampleName = "X121270372616437_1_1_1_1", ColorBySampleSource = "---", 
    ColorBySampleDynamicSource = FALSE, ShapeBy = "None", SizeBy = "None", 
    SelectionAlpha = 0.1, ZoomData = numeric(0), BrushData = list(), 
    VisualBoxOpen = FALSE, VisualChoices = "Color", ContourAdd = FALSE, 
    ContourColor = "#0000FF", PointSize = 1, PointAlpha = 1, 
    Downsample = FALSE, DownsampleResolution = 200, CustomLabels = FALSE, 
    CustomLabelsText = "X121270372616437_1_1_1_1", FontSize = 1, 
    LegendPointSize = 1, LegendPosition = "Bottom", HoverInfo = TRUE, 
    LabelCenters = FALSE, LabelCentersBy = "gender", LabelCentersColor = "#000000", 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    DataBoxOpen = FALSE, RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())

################################################################################
# Settings for Complex heatmap 2
################################################################################

initial[["ComplexHeatmapPlot2"]] <- new("ComplexHeatmapPlot", Assay = "logcounts", CustomRows = TRUE, 
    CustomRowsText = "LAG3\nPDCD1\nCTLA4\nC1QC\nSPP1", ClusterRows = FALSE, 
    ClusterRowsDistance = "spearman", ClusterRowsMethod = "ward.D2", 
    DataBoxOpen = FALSE, VisualChoices = "Annotations", ColumnData = "cluster_kmeans_k6", 
    RowData = character(0), CustomBounds = FALSE, LowerBound = NA_real_, 
    UpperBound = NA_real_, AssayCenterRows = FALSE, AssayScaleRows = FALSE, 
    DivergentColormap = "purple < black < yellow", ShowDimNames = "Rows", 
    LegendPosition = "Bottom", LegendDirection = "Horizontal", 
    VisualBoxOpen = FALSE, NamesRowFontSize = 10, NamesColumnFontSize = 10, 
    ShowColumnSelection = TRUE, OrderColumnSelection = TRUE, 
    VersionInfo = list(iSEE = structure(list(c(2L, 5L, 2L)), class = c("package_version", 
    "numeric_version"))), PanelId = 2L, PanelHeight = 500L, PanelWidth = 4L, 
    SelectionBoxOpen = FALSE, RowSelectionSource = "---", ColumnSelectionSource = "---", 
    RowSelectionDynamicSource = FALSE, ColumnSelectionDynamicSource = FALSE, 
    RowSelectionRestrict = FALSE, ColumnSelectionRestrict = FALSE, 
    SelectionHistory = list())
```


## Figures figures figures

Do some by hand, then extract the code

```{r}

```

## Adding a tour to tell things to people - they like to hear or read about the data!

```{r}

```


library(zellkonverter)

sce_tica <- readH5AD("TICAtlas_downsampled_1000.h5ad")


## Fully fledged instance of iSEE

```{r}
iSEE(sce, initial = initial, appTitle = "iSEE ❤️ TICA")
```

# Session info

```{r}
sessionInfo()
```



